A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-10-14 draft.tex
Not a challenge
Fixes #22
Introduction
Proof read and deleted a repetiive paragraph. @JJ  please check.

Besides being all population-based, they also share the common characteristic
of creating an initial set of random candidate solutions that are later used to
generate a new set of candidates, using a nature-inspired heuristic.
Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be (in general) evaluated independently of 
others, and similarly, each population could evolve in isolation. Researchers use the term
multi-population based methods when generally referring to techniques using
many populations as part of their optimization strategy. 
Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these
algorithms that adds to this simple parallel execution some form of
interaction between the independently running populations.
The island model was one of the first techniques proposed for parallelization,
which led to an early proof of speedup   \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}.

% Probably it's not the time for this, but maybe it would be interesting to 
% mention that asynchronous implementations in most cases need a change in 
% the underlying algorithm, while synchronous might not need it - JJ 
% TO DO - Mario 
Moreover, parallel implementations can run in two different ways: synchronous and
asynchronous. When operations run in parallel but every node must maintain synchrony with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  master/worker model, the master needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other, so there's no time lost waiting for other populations to
arrive to the synchronization point; results arriving from other workers are 
incorporated as soon as they are received. Following the previous example, now the master operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

The trend concerning the parallel execution multi-population algorithms goes from
earlier hardware-based implementations using transputers
\cite{gorges1990explicit} through 
multi-core systems \cite{Serrano2008,lai2019adaptive} to multi-threaded \cite{merelo2019scaling} ones.
And recently the focus has been on exploiting a higher number of processing
units by using GPUs \cite{tan2015survey,li2007efficient}, or distributed
systems, including web-based \cite{JSON} systems,
map-reduce \cite{fazenda2012} implementations,  grids \cite{munawar2010design,Gonzalez09},
voluntary computing systems \cite{MilkyWay,merelo2016nodio}, 
and, more importantly, cloud computing
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. Cloud computing has become the standard way of running
enterprise applications, not only because of the convenience of the
pay-as-you-go model or the non-existent sunk cost of physical facilities, but also because
it offers a way of describing the infrastructure as part of the code, so that it
is much easier to reproduce results and this has been a boon for scientific
computing.  However,  cloud computing has also been evolving, going from simply
putting on virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
built on a single stack of services that communicate by layers, to microservices
\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} that favor the parallel and asynchronous communication of
heterogeneous resources; in the process, a new methodology for designing {\em cloud
native applications} has been created, with brand new methods and
techniques \cite{Baldini2016287}. In these architectures, services or even
functions are seen as independent processing nodes, departing from the monolithic
and even distributed paradigms, and becoming a loosely coupled collection of {\em
stateless functions} \cite{malawski2017serverless}, that react to events and
only come into existence while they are doing some processing.
Systems developed with the patterns outlined above have the
properties of a reactive system \cite{boner2014reactive} and they are commonly
more flexible, fault-tolerant, and scalable. % Some properties and reference - Mario  
% Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ

@@

    \item To scale the system, when it is possible, there are several options, including the  use of stateless functions for
    processing.

@@

% I have moved these sentences here, since they seem to fit the best.(OK - M)
Beyond the choice of design methodology and implementation, researchers need to
consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
the interaction between them, the search area of each population, and the search
strategy and parametrization of each population. We must take into account
these high-level issues when designing a parallel architecture.

Taking these factors into account, in this work, we present EvoSwarm, 
a container-based application that reactively processes
isolated and
heterogeneous populations. The platform uses the {\tt docker-compose} tool for
defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in
\footnote{\url{https://github.com/mariosky/EvoSwarm}}.
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

Docker container images host the stateless functions responsible for 
running a search strategy on populations they receive as messages. 
A single container can pull many messages on its lifetime, processing one at a time. 
% clarify "on several populations" Sequentially? At the same time? - JJ 
% done - M
These functions could also be ported
directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

This paper extends our earlier publication on the topic
\cite{guervos2018introducing}, and we highlight the main contributions as
follows:
We have proved that cloud-native applications are
able to easily accommodate multi-paradigm bioinspired algorithms
with little efforts of parametrization and
that a combination of algorithms has better speedup and performance
than any single one taken independently.

@@

% Maybe here you should make reference to the sections where you are actually doing it (after conveniently rephrasing it) - JJ
% These highlights go in the web page of the paper, they need to be short and to the point, 
% I mention them briefly when presenting the paper's sections - M
\begin{itemize}
    \item First, we present the design and implementation of a reactive 
    container-based application for the asynchronous execution of multi-population 
    algorithms. The source code and example container definitions are
    publicly available. % Not really a contribution. It's the mean to
    % achieve the contribution - JJ
    % Maybe you should check this out again ... - JJ
    \item Second, we propose a new method for the deployment and execution of 
    multiple experiments by specifying the infrastructure as part of an 
    experiment definition in both local or cloud environments,
    facilitating the reproduction of experimental results. % Not clear
                                % what the contribution is. If this
                                % has been used for the first time in
                                % an EA environment, it should be
                                % stated that way. If not, it's also a
                                % mean to achieve the contribution - JJ
    \item Third, the application is compatible with the COCO benchmark 
    framework, allowing researchers to compare the performance of their 
    algorithms with other works. % Not really a contribution, but an
                                % implementation detail. Convenient,
                                % but does not advance the state of
                                % the art - JJ
    \item Fourth, we present an empirical study to validate the 
    applicability of our application,  measuring the execution time and  
    speedup.

    \item Lastly, we present a comparison between homogeneous and an ensemble of 
    multi-populations, using Genetic Algorithms (GAs) and Particle 
    Swarm Optimization (PSO) in a benchmark for the optimization of 
    continuous functions. % "Validate the applicability" falls short
                          % of being also a contribution. It must
                          % advance the state of the art, in which way
                          % does it advance? Speed (needs to be
                          % proved) Convenience (needs to be proved
                          % too) Scalability? (needs to be proved
                          % also). - JJ 
\end{itemize}

The organization of the paper is as follows: First, in Section \ref{multi}, we present a
background of the fundamental issues of integrating nature-inspired optimization
and multi-population methods. Section \ref{soa} presents state of the art relevant to
our work. In Section \ref{method}, we present the proposed method and the container-based
application in Section \ref{docker} and the workflow for reproducible experiment design in Section \ref{experiment_flow}. % No reference to section 6.
Section \ref{setup} describes the design of the empirical evaluation to assess the effectiveness
of the method,  with respect to scalability and % This is a subsection, fix it. 
the performance of heterogeneous populations.  % No references to sections 9 and 10
Finally, we offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.

@@

Later approaches use microservices \cite{khalloof2018generic}, or were
systematized \cite{salza2019speed} achieving speed-up through the
creation of an abstraction layer over the evolutionary algorithm
services, and also automation of the whole workflow. This research
bumped into some problems, since speedup is limited, mainly due to the
non-automatic scaling of services and also the presence of a single
master which is the one that runs the evolutionary
algorithms. However, this master-worker architecture can be turned
around with other kind of evolutionary architectures, such
as pool-based systems
\cite{valenzuela2015implementing,merelo2012sofea,sofea:cec2012} where
a better match to the inherently heterogeneous nature of cloud-based
architectures. Pool-based systems are closer to serverless systems
\cite{malawski2017serverless}, and they pull from the pool and run
whole algorithms, instead of just evaluation. The pool from where they
pull is the common store of results, and can be accessed in an
asynchronous way, which suits better the nature of cloud systems.

@@

In this experiment, we evaluate how the number of populations and the number of
workers in the system are related to the scalability of the system.
In the next section, we describe the experimental setup, and a discussion of the results
follows, with respect to the speedup in terms of evaluations per second (Section~\ref{sec:speedup}), and 
execution time (Section~\ref{sec:exec-time}).

@@

\subsubsection{Results Regarding Speedup}
\label{sec:speedup}

@@

When comparing the results in this section, we performed the Wilcoxon Signed-Rank test, with a
significance level of 5\%. We indicate with an asterisk those results that are
not significant throughout the tables of this section.  % Where? - JJ
% Answered - Mario
To measure the speedup obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work, because we are using a stochastic search. 
We are going to measure the total increase of work done in a second by adding more workers.

\begin{table}[h!tbp]
  \caption{Speedup by worker and dimension, taking one worker as the
    baseline. Speedup is obtained by dividing the evaluation ratio of the baseline by the 
    ratio of the increased number of workers.}
% Clarify here what you are measuring. Number of evaluations for
% whatever divided by whatever - JJ
% Done - Mario
  \label{tab:speedup-table}
  \vspace{0.25cm}
  \centering

@@

%
As expected, within the same number of populations, there is a linear speedup
when we add more workers. The speedup is marginal in those cases where there 
are fewer populations than workers.  The reason for this is that at any given time, 
there are at least two populations in the message queue, leaving some of the workers 
waiting for populations to be available in the queue.

\begin{table}[h!tbp]
  \small
  \caption{Speedup achieved with increasing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. 
    Best and Worst are highlighted (bold and underlined)}
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{llllll}
  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
  & 10 & 8w & 16w & 1.05 & 0.072660 \\
  & 20 & 1w & 2w  & \textbf{2.09} & 1.51E-11 \\
  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
  & 20 & 8w & 16w & {\ul 1.03} & 0.214482 \\
  &    &    &     &      &          \\
10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
  &    &    &     &      &          \\
 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
 \end{tabular}
  \end{table}

Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
one worker. We achieved the maximum speedup when using four workers or more with
twenty populations. On the other hand, the maximum speedup when we consider the
cost it was with eight workers. The best speedups for each number of workers are
shown in bold in the table.

Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
the previous configuration, for example, when increasing from four to eight
workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
increment in all cases. Again, the best effect is achieved when the number of
populations is higher than the number of workers. Configurations following this
basic rule doubled the speed or where near it. Using five populations, we
obtained the best and worst speedup depending on the number of workers (shown in
underline and bold, respectively).

@@

In the previous section, we have shown that as we increase the number of
populations and workers, we increase the ratio of the number of evaluations per
second in the same proportion. However, it is vital to notice that we can also
affect the number of evaluations we need to execute to find a solution. This
reduction means that the multi-population based algorithm can perform better
when we increase the number of populations, or in some cases, even workers, as the experiments
show. If we increase the evaluations per second and, at the same time, decrease
the number of evaluations needed, we can have a superlinear speedup of the
algorithm, if we measure the time required to complete an instance.

@@

In Table~\ref{tab:time} we can see that in some cases we have a supralinear
speedup, we highlighted these in boldface and indicated the speedup value in
parenthesis. The best overall times are underlined. As expected, increasing the
number of workers for a fixed dimension and populations reduced the time
required to find a solution, as long as the population number was higher than
the number of workers. The values that had no statistical difference (in the
Wilcoxon test) when increasing the number of workers have an asterisk.

>>>>

Besides being all population-based, they also share the common characteristic of 
creating an initial set of random candidate solutions later used by a 
nature-inspired heuristic to generate a new set of candidates. 
Population-based algorithms are intrinsically parallel.
The fitness of each individual can be (in general) evaluated independently of 
others, and similarly, each population could evolve in isolation. Researchers use the term
multi-population based methods when referring to techniques using
many populations as part of their optimization strategy. 
Since earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase these algorithms' scalability
by adding some form of interaction between populations running in isolation.
The island model was one of the first techniques proposed for parallelization,
which led to early proofs of speed-up \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the idea. 
Researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}.

% Probably it's not the time for this, but maybe it would be interesting to 
% mention that asynchronous implementations in most cases need a change in 
% the underlying algorithm, while synchronous might not need it - JJ 
% TO DO - Mario 
Moreover, parallel implementations can run in two different ways: synchronous and
asynchronous. When operations run in parallel, but every node must maintain synchrony with
the other operations, they all need to finish before moving to the next
step. For instance, in a  master/worker model, masters need to wait for all
workers to complete their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other, so there's no time lost waiting for other populations to
arrive at the synchronization point; results coming from other workers are 
incorporated as soon as they are received. Following the previous example, now the master operation can
continue to the next iteration, even if a worker has not finished its work.  We can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}. In this work, we follow this
practice.

The trend concerning the parallel execution multi-population algorithms goes from
earlier hardware-based implementations using transputers
\cite{gorges1990explicit} through 
multi-core systems \cite{Serrano2008,lai2019adaptive} to multi-threaded \cite{merelo2019scaling} ones.
Moreover, recently, the focus has been on exploiting a higher number of processing
units by using GPUs \cite{tan2015survey,li2007efficient}, or distributed
systems, including web-based \cite{JSON} systems,
map-reduce \cite{fazenda2012} implementations,  grids \cite{munawar2010design,Gonzalez09},
voluntary computing systems \cite{MilkyWay,merelo2016nodio}, 
and, more importantly, cloud computing
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}.

%% JJ Can you improve the redability of the paragraph below? 
%% Grammarly suggests that is hard to read, verly long sentences.
Cloud computing has become the standard way of running
enterprise applications.  not only because of the convenience of the
pay-as-you-go model or the non-existent sunk cost of physical facilities, but also because
it offers a way of describing the infrastructure as part of the code, so that it
is much easier to reproduce results and this has been a boon for scientific
computing.  However,  cloud computing has also been evolving, going from simply
putting on virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
built on a single stack of services that communicate by layers, to microservices
\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} 
that favor the parallel and asynchronous communication of
heterogeneous resources; in the process, a new methodology for designing 
{\em cloud-native applications} has been created, with brand new methods and
techniques \cite{Baldini2016287}. 
In these architectures, services or even
functions are seen as independent processing nodes, departing from the monolithic
and even distributed paradigms, and becoming a loosely coupled collection of {\em
stateless functions} \cite{malawski2017serverless}, that react to events and
only come into existence while they are doing some processing.
Systems developed with the patterns outlined above have the
properties of a reactive system \cite{boner2014reactive}, and they are commonly
more flexible, fault-tolerant, and scalable. % Some properties and reference - Mario  
% Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ

@@

  %  \item To scale the system, when it is possible, there are several options, including the  use of stateless functions for
  %  processing.

@@

% I have moved these sentences here, since they seem to fit the best.(OK - M)
Beyond the choice of design methodology and implementation, researchers need to
consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
the interaction between them, the search area of each population, and the search
strategy and parametrization for each population. We must take into account
these high-level issues when designing a parallel architecture.

Considering these factors, in this work, we present EvoSwarm, 
a container-based application that reactively processes
isolated and heterogeneous populations. The platform uses the {\tt docker-compose} tool for
defining and running experiments as multi-container applications. EvoSwarm is also released with a free license
and available in a public 
GitHub repository
\footnote{\url{https://github.com/mariosky/EvoSwarm}}.
We preferred the use of a container-based implementation for this prototype because
it is more suitable for development and replication without needing
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.
%% Here we could talk about Cloud-based platform for executing the application
%% and the cloud-lock risk - Mario

%% This is explained later in detail, we do not needed here.
%% JJ Please check this - Mario

% Docker container images host the stateless functions responsible for 
% running a search strategy on populations they receive as messages. 
% A single container can pull many messages in its lifetime, processing one at a time. 
% % clarify "on several populations" Sequentially? At the same time? - JJ 
% % done - M
% These functions could also be ported
% directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% % Need to improve the next paragraph or maybe move it to another place - Mario
% Researchers then define the configuration of services used by
% a multi-population algorithm using a YAML file. Any researcher can deploy and
% start all the services from this configuration file with a single command. Once
% the services are up, multiple instances of an experiment can be sent to a queue
% for their execution. The platform includes containerized services that are
% compatible with other container orchestration technologies like Kubernetes and
% Docker Swarm. The services included are a Redis Server used for the Message
% Queue implementation and logging and an Experiment Controller implemented in
% Python. Developers of population processing functions have the freedom to
% implement new population-based algorithms or operators in the language they
% choose inside a container, as they only need to subscribe to a channel of the
% Message Queue and have the ability to read the population and parameters as JSON  
% file.

This paper extends our earlier publication on the topic
\cite{guervos2018introducing}, and we highlight the main contributions as
follows:
We have proved that cloud-native applications can easily accommodate multi-paradigm bioinspired algorithms
with little efforts of parametrization and
that a combination of algorithms has better speed-up and performance
than any single one taken independently.

@@

% Maybe here you should make reference to the sections where you are actually doing it (after conveniently rephrasing it) - JJ
% These highlights go in the web page of the paper, they need to be short and to the point, 
% I mention them briefly when presenting the paper's sections - M
\begin{itemize}
    \item First, we present the design and implementation of a reactive 
    container-based application for the asynchronous execution of multi-population 
    algorithms. The source code and example container definitions are
    publicly available. % Not really a contribution. It's the mean to
    % achieve the contribution - JJ
    % Maybe you should check this out again ... - JJ
    \item Second, we propose a new method for the deployment and execution of 
    multiple experiments by specifying the infrastructure as part of an 
    experiment definition in both local or cloud environments,
    facilitating the reproduction of experimental results. % Not clear
                                % what the contribution is. If this
                                % has been used for the first time in
                                % an EA environment, it should be
                                % stated that way. If not, it's also a
                                % mean to achieve the contribution - JJ
    \item Third, the application is compatible with the COCO benchmark 
    framework, allowing researchers to compare their 
    algorithms' performance against other works. % Not really a contribution, but an
                                % implementation detail. Convenient,
                                % but does not advance the state of
                                % the art - JJ
    \item Fourth, we present an empirical study to validate 
    our application's applicability, measuring the execution time and speed-up.

    \item Lastly, we present a comparison between homogeneous and an ensemble of 
    multi-populations, using Genetic Algorithms (GAs) and Particle 
    Swarm Optimization (PSO) using a benchmark for 
    continuous function optimization. % "Validate the applicability" falls short
                          % of being also a contribution. It must
                          % advance the state of the art, in which way
                          % does it advance? Speed (needs to be
                          % proved) Convenience (needs to be proved
                          % too) Scalability? (needs to be proved
                          % also). - JJ 
\end{itemize}

We organized the paper as follows: First, in Section \ref{multi}, we present a
background of the fundamental issues of integrating nature-inspired optimization
and multi-population methods. Section \ref{soa} presents state of the art relevant to
our work. In Section \ref{method}, we present the proposed method and the container-based
application in Section \ref{docker} and the workflow for reproducible experiment design in Section \ref{experiment_flow}. % No reference to section 6.
Section \ref{setup} describes the design of the empirical evaluation to assess the effectiveness
of the method,  with respect to scalability and % This is a subsection, fix it. 
the heterogeneous populations' performance.  % No references to sections 9 and 10
Finally, we offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.

@@

Later approaches use microservices \cite{khalloof2018generic}, or were
systematized \cite{salza2019speed} achieving speed-up through the
creation of an abstraction layer over the evolutionary algorithm
services, and also automation of the whole workflow. This research
bumped into some problems, since speed-up is limited, mainly due to the
non-automatic scaling of services and also the presence of a single
master which is the one that runs the evolutionary
algorithms. However, this master-worker architecture can be turned
around with other kind of evolutionary architectures, such
as pool-based systems
\cite{valenzuela2015implementing,merelo2012sofea,sofea:cec2012} where
a better match to the inherently heterogeneous nature of cloud-based
architectures. Pool-based systems are closer to serverless systems
\cite{malawski2017serverless}, and they pull from the pool and run
whole algorithms, instead of just evaluation. The pool from where they
pull is the common store of results, and can be accessed in an
asynchronous way, which suits better the nature of cloud systems.

@@

In this experiment, we evaluate how the number of populations and the number of
workers in the system are related to the scalability of the system.
In the next section, we describe the experimental setup, and a discussion of the results
follows, with respect to the speed-up in terms of evaluations per second (Section~\ref{sec:speedup}), and 
execution time (Section~\ref{sec:exec-time}).

@@

\subsubsection{Results Regarding Speed-up}
\label{sec:speedup}

@@

When comparing the results in this section, we performed the Wilcoxon Signed-Rank test, with a
significance level of 5\%. We indicate with an asterisk those results that are
not significant throughout the tables of this section.  % Where? - JJ
% Answered - Mario
To measure the speed-up obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work, because we are using a stochastic search. 
We are going to measure the total increase of work done in a second by adding more workers.

\begin{table}[h!tbp]
  \caption{Speed-up by worker and dimension, taking one worker as the
    baseline. Speed-up is obtained by dividing the evaluation ratio of the baseline by the 
    ratio of the increased number of workers.}
% Clarify here what you are measuring. Number of evaluations for
% whatever divided by whatever - JJ
% Done - Mario
  \label{tab:speedup-table}
  \vspace{0.25cm}
  \centering

@@

%
As expected, within the same number of populations, there is a linear speed-up
when we add more workers. The speed-up is marginal in those cases where there 
are fewer populations than workers.  The reason for this is that at any given time, 
there are at least two populations in the message queue, leaving some of the workers 
waiting for populations to be available in the queue.

\begin{table}[h!tbp]
  \small
  \caption{Speed-up achieved with increasing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. 
    Best and Worst are highlighted (bold and underlined)}
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{llllll}
  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speed-up            & p-value\\
  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
  & 10 & 8w & 16w & 1.05 & 0.072660 \\
  & 20 & 1w & 2w  & \textbf{2.09} & 1.51E-11 \\
  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
  & 20 & 8w & 16w & {\ul 1.03} & 0.214482 \\
  &    &    &     &      &          \\
10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
  &    &    &     &      &          \\
 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
 \end{tabular}
  \end{table}

Table~\ref{tab:speedup-table} shows the speed-up obtained against the baseline of
one worker. We achieved the maximum speed-up when using four workers or more with
twenty populations. On the other hand, the maximum speed-up when we consider the
cost it was with eight workers. The best speed-ups for each number of workers are
shown in bold in the table.

Table~\ref{tab:speedup:test} shows the speed-up obtained taking as the baseline
the previous configuration, for example, when increasing from four to eight
workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
increment in all cases. Again, the best effect is achieved when the number of
populations is higher than the number of workers. Configurations following this
basic rule doubled the speed or where near it. Using five populations, we
obtained the best and worst speed-up depending on the number of workers (shown in
underline and bold, respectively).

@@

In the previous section, we have shown that as we increase the number of
populations and workers, we increase the ratio of the number of evaluations per
second in the same proportion. However, it is vital to notice that we can also
affect the number of evaluations we need to execute to find a solution. This
reduction means that the multi-population based algorithm can perform better
when we increase the number of populations, or in some cases, even workers, as the experiments
show. If we increase the evaluations per second and, at the same time, decrease
the number of evaluations needed, we can have a superlinear speed-up of the
algorithm, if we measure the time required to complete an instance.

@@

In Table~\ref{tab:time} we can see that in some cases we have a supralinear
speed-up, we highlighted these in boldface and indicated the speed-up value in
parenthesis. The best overall times are underlined. As expected, increasing the
number of workers for a fixed dimension and populations reduced the time
required to find a solution, as long as the population number was higher than
the number of workers. The values that had no statistical difference (in the
Wilcoxon test) when increasing the number of workers have an asterisk.
